---
title: "a3_task3_ali_martin"
author: "Ali Martin"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Read in the urchins data

```{r}
urchins <- read.csv(here("data", "urchins.csv")) %>% clean_names()

```


## Part 1: Size comparisons of purple and red urchins

### Data Visualization

Make a graph comparing sizes of purple and red urchins 
```{r}
ggplot(data = urchins,
       aes(x = size)) +
  geom_histogram(bins = 12, aes(fill = common_name)) +
  scale_fill_manual(values = c("darkmagenta", "red2")) +
theme_minimal() + 
  labs(x = "Size (cm)",
       y = "Number of Urchins",
       title = "Size Distribution of Purple and Red Urchins in Santa Barbara",
       fill = "Key") +
  facet_wrap(~ common_name)
        
         
  
         
```

**Figure 1:** The above graph shows the size comparison of red and purple urchins observed at 5 different sites along the Santa Barbara coast. 

### Descriptive Statistics Table

```{r}
urchins_stats <- urchins %>% #create descriptive stats table from urchins data
  group_by(common_name) %>% #group by common name
  summarize( #form a table
    mean_urchins = round(mean(size), 2), #calculate mean and round sig figs
    median_urchins = median(size), #calculate median and round sig figs
    sd_urchins = round(sd(size), 2), #calculate std deviation and round sig figs
    sample_size = n() #calculate sample size
  )

kable(urchins_stats, #create stats table usinc kablextra
      caption ="Descriptive Statistics for purple and red urchin sizes", #form caption
      col.names = c("Type of Urchin", #update column names
                    "mean urchin size (cm)",
                    "median urchin size (cm)",
                    "Standard deviation of urchin size (cm)",
                    "sample size")) %>% 
  kable_styling(full_width=F)

```

Based on the histogram above, it is clear that the distribution of purple urchins is positively skewed, and the distribution of red urchins is closer to a normal distribution. The spread of purple urchins is less than the spread of the red urchins, but the purple urchins have a much higher median value than the red urchins.

### Section 2: Explore purple urchin counts by size and year

```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% #filter to include only purple urchins
  mutate(date = mdy(date)) %>% #change date to class "date"
  mutate (year = year(date)) %>%
  mutate(site_full = case_when(
    site =="NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))
  
  
```

Store a table containing total number of purple urchins at each site by year

```{r}
purple_urchins %>% 
  group_by(common_name, year, site_full) %>% 
  summarize()
    
  
```

Make a finalized graph of counts of purple urchins recorded at each site per year

