---
title: "Task 4: Naples Reef Purple Urchins"
author: "Ali Martin"
date: "10/28/2021"
output: html_document
---
## Setup

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message= FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(kableExtra)
```

## Read in the data

```{r}
urchins_data <- read_csv(here("data", "urchins.csv")) %>% clean_names()
```

## Data Wrangling

```{r}
naples_urchins <-urchins_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  group_by(treatment) %>% 
  mutate(date= mdy(date)) %>%  #convert date to class "date"
  mutate(year = year(date)) %>%  #create new column as year
  mutate(site_full = case_when( #add in the full site names
    site =="NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef")) %>% 
  filter(year == "2010") #filter to only include data from 2010
  
```

## Data Visualization (Histogram and QQ Plot):


```{r}
ggplot(data = naples_urchins, #plot data from naples urchins
       aes(x = size)) + #assign x value to size of urchin
  geom_histogram(bins = 12, aes(fill = treatment)) + #plot histogram with 12 bins organized by treatment
  scale_fill_manual (values = c("blue", "green")) + #assign color values and...
  theme_classic() + #assign theme
  labs(x = "size (cm)", y = "Number of Urchins", title = "Figure 1: Size Distribution of Naples Urchins By Treatment") + #assign labels
  facet_wrap(~ treatment) #separate plots by treatment

ggplot(data = naples_urchins, aes(sample = size)) + #plot 
  geom_qq() + #make QQ plot
  geom_qq_line() + #add qq line to show linear relationship
  labs(title = "Figure 2: QQ Plot of Naples Urchin Size Distribution by Treatment") + #add title
  facet_wrap(~treatment) #wrap to separate graphs by treatment
```

The above graphs shows that there is not a significant difference between the two distributions. On the histogram, both the Control and Annual graphs show a loose postively skewed distribution. The only significant difference I've noticed is that in the control, there are over 600 urchins that are within a size of around 5 cm, where as in the annual plot, there are less than 600 urchins with size ~5 cm .

### Summary Table

```{r}

naples_urchins_stats <- naples_urchins %>% #create descriptive stats table from naples urchins data
  group_by(treatment) %>% #group by treatment
  summarize( #form a table
    mean_naples_urchins = round(mean(size), 2), #calculate mean and round sig figs
    median_naples_urchins = median(size), #calculate median and round sig figs
    sd_naples_urchins = round(sd(size), 2), #calculate std deviation and round sig figs
    sample_size = n() #calculate sample size
  )

kable(naples_urchins_stats, #create stats table using kablextra
      caption ="Figure 3: Descriptive Statistics for purple Naples urchins for Annual and Control treatments", #form caption
      col.names = c("Treatment", #update column names
                    "mean urchin size (cm)",
                    "median urchin size (cm)",
                    "Standard deviation of urchin size (cm)",
                    "sample size")) %>% 
  kable_styling(full_width=F)

```

The summary table shows that the mean urchin size of both annual and control treatments is only off by .01. The median urchin size is exactly the same (4.5 cm) for both treatments, and the standard deviation is only off by .07, which is still very close as well. 
